Title goes here

### 2016-10-23 paper/paper.tex
asd

The number of flying airplane has increased much in the last decade. This led to
many problems, mostly related to security and air traffic controls. It was clear
that radar surveillance systems would not be sufficient to monitor flights,
since airplanes are passive entity. With the introduction of new technologies, a
new standard called ADS-B has been developed to actively monitor aircrafts and
manage routings.

\textit{Automatic Dependant Surveillance - Broadcast} (ADS-B) is the last
generation standard for air traffic monitoring. The main idea of this technology
is that each aircraft broadcast its position and status using internal sensors,
without relying on external entities. On the other side, each airplane is able
to receive the messages and detect the position of other vehicle autonomously,
thus maintening a minimum distance and avoiding collision.

ADS-B messages are sent without any encryption and can be received and decoded
by everyone. Morover this equipment will be mandatory in 2017 in most countries.
We obtained a dataset, provided by OpenSky network, containing all the ADS-B
messages sent by airplanes and collected by volunteers. The dataset includes
data about flights from the 18\textsuperscript{th} September 2016 to the
24\textsuperscript{th} September 2016 and collected by about 70 sensors spread
across Europe.

In this paper we explain our analysis of such dataset, starting from the
decoding of the position. The main goals of this project are:
\begin{itemize}
  \item Detecting the standard routes between all the pairs of airports;
  \item Detecting abnormal routes between pairs of airports;
  \item Calculate the wastage of fuel travelling through the standard route
  instead of flying over a straight line.
\end{itemize}

We will show that, though most standard routes are really close to the straight
line, there are some routes that are clearly away from it. Moreover, for some
couples of airports, there are few routes that differ from the standard one.

\section{Related Work}
% meaningfully summarize related literature to explain what this project builds
% on, in particular, discuss the two research papers chosen with this project
% (and discussed in your presentation)

@@

\section{Research Questions}
% which questions is this project trying to answer, and/or hypotheses to
% investigate? These should cover both the project topic, as well as the
% technological side (aptness of the tools for the job, scalability of the
% solution).

@@

Analyzing the entire dataset require a big computational power, therefore use a
cluster to perform analyze the data. However, before running on the cluster, we
tried the algorithms locally on a small subset of data to test correctness end
efficiency and to obtain statistics. For this reason, we used different
technologies depending on where we ran the analysis.

On the cluster, we chose to use \textit{Apache Spark} because it is extremely
fast while providing an easy interface, resulting suitable for our application.
In particular, the \textit{Scala} version was chosen instead of PySpark because
of the strongly typing and the functional programming approach. In addition to
this, the Java library \textit{libadsb} was used to decode the position
messages.

Locally, in addition to \textit{Scala} we used \textit{Haskell} and
\textit{Python} as scripting languages. The first one was suitable for data
manipulation and algorithms testing, while the second one offers a convenient
way to plot data using the library \textit{Matplotlib}.

@@

\paragraph{Flight detection}

@@

\paragraph{Airlines}

@@

\paragraph{Diverging flights}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.35]{imgs/schiphol-linate.png}
  \caption{Results of standard (blue) and diverging (red) routes detection for
    flights from Amsterdam Schiphol to Milano Linate. It can be seen that the
    diverging route on the right is very similar to the standard route, but the
    clustering algorithm fails to classify it as such.}
  \label{fig:schiphol-linate}
\end{figure}

@@

The results from the execution of the route detection algorithm are mixed.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete.
Of 1165 of the total combinations of airports, about 85 \% of them has their
standard route determined by under 10 flights. These pairs usually correspond
to minor airports, and the accuracy of the computed standard route has
necessarily to be considered pretty low. For major airports, however, the
available data rises significantly. As an example, the standard route between
Schiphol and London Gatwick and converse has been determined by 39,
respectively 42 routes.

>>>>

The number of flying airplanes has increased much in the last decade. This led to
many problems, mostly related to security and air traffic control. It was clear
that radar surveillance systems would not be sufficient to monitor flights,
since airplanes are passive entities. With the introduction of new technologies, a
new standard called ADS-B has been developed to actively monitor aircrafts and
manage routings.

\textit{Automatic Dependant Surveillance - Broadcast} (ADS-B) is the last
generation standard for air traffic monitoring. The main idea of this technology
is that each aircraft broadcasts its position and status using internal sensors,
without relying on external entities. On the other hand, each airplane is able
to receive the messages and detect the position of other vehicles autonomously,
thus allowing it to maintain a minimum distance and avoid collision.

ADS-B messages are sent without any encryption and can be received and decoded
by anyone. We were able to obtain a dataset, provided by OpenSky network,
containing all the ADS-B messages sent by airplanes and collected by volunteers
during a time period spanning a week. The dataset includes data from September
18\textsuperscript{th} 2016 to September 24\textsuperscript{th} 2016, as
collected by about 70 sensors spread across western Europe. It is the purpose of
this paper to explain our approach in analyzing and extracting data from such
dataset. It is particularly intresting to see to which point data like the one
offered by OpenSky is suitable to get useful information about flights. In this
project, we concentrate on analyzing the air traffic and its patterns, and in
particular why try to discover standard routes between pairs of
airports, as well as flights that diverge from them.

\section{Related Work}

@@

\section{Research Questions}

@@

Analyzing the entire dataset requires a lot of computational power, hence the
need to use a cluster to perform analysis on the data. However, before executing
on the cluster, it is essential to be able to try the algorithms locally on a
small subset of data, in order to test their correctness end efficiency, and to
obtain useful statistics. For this reason, we used different technologies
depending on where we ran the analysis.

On the cluster, we chose to use \textit{Apache Spark} because it is extremely
fast while providing an easy and convenient development interface. In
particular, the \textit{Scala} version was chosen instead of PySpark because of
the strongly typing and the functional programming approach. In addition to
this, the Java library \textit{libadsb} was used to decode the raw ADS-B
messages.

On our local datasets, in addition to \textit{Scala} we used \textit{Haskell}
and \textit{Python} as scripting languages for fast prototyping. The first one
was suitable for data manipulation and algorithms testing, while the second one
offers a convenient way to plot data using the library \textit{Matplotlib}.

@@

\subsection{Flight detection}

@@

\subsection{Airlines}

@@

\subsection{Diverging flights}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.35]{imgs/schiphol-linate.png}
  \caption{Results of standard (blue) and diverging (red) routes detection for
    flights from Amsterdam Schiphol to Milano Linate, during our previous
    experiments. It can be seen that the diverging route on the right is very
    similar to the standard route, but the clustering algorithm fails to
    classify it as such. This made us increase the threshold in the
    classification of routes in clusters.}
  \label{fig:schiphol-linate}
\end{figure}

@@

The results from the execution of the route detection algorithm are quite good.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete. Of 1165 of
the total combinations of airports, about 85 \% of them has their standard route
determined by under 10 flights. These pairs usually correspond to minor
airports, and the accuracy of the computed standard route has necessarily to be
considered pretty low. For major airports, however, the available data rises
significantly. As an example, the standard route between Schiphol and London
Gatwick and converse has been determined by 39, respectively 42, routes. This
gives more confidence on the accuracy of the result.
